Language-Based Security for Low-Level MPC

### 2024-07-15 ppdp24/model.tex
checkpoint

We define joint probability mass functions (pmfs) in the standard
manner, though following \cite{barthe2019probabilistic} we use
memories to denote mappings of variables to values (i.e., outcomes),
so for example given a pmf $\pmf$ we will write $\pmf(\{ \elab{\secret{x}}{1}
\mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to denote the (joint) probability that
$\elab{\secret{x}}{1} = 0 \wedge \elab{\mesg{y}}{2} = 1$.
\begin{definition}
  A \emph{probability mass function} $\pmf$ is a function
  mapping memories in $\mems(X)$ for some $X$ to values in $\mathbb{F}_p$, such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ the \emph{marginal distribution} of variables $X$
  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
  $$
  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
  $$
\end{definition}

\begin{definition}
  Given $\margd{\pmf}{X}$, let $\stores$ be a set of memories with the
  same domain $Y \subseteq X$. Then the \emph{conditional distribution given
  $\stores$}  denoted
  $\condd{\pmf}{X}{\stores}$ is a distribution with domain $X$ where for all
  $\store \in \mems(X)$:
  $$
  \condd{\pmf}{X}{\stores}(\store) =
  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
  $$
  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
\end{definition}
To recover familiar notation we allow the syntactic
sugarings $\condd{\pmf}{X}{\store}  \defeq \condd{\pmf}{X}{\{ \store\}}$, and
$\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$ and $\pmf(\store|\stores) \defeq
\condd{\pmf}{X}{\stores}(\store)$ where $\dom(\store) = X$.
%\begin{eqnarray*}
%  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
%  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
%  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
%\end{eqnarray*}

We also define the \emph{support} of a distribution in the usual manner-
it is the set of values a set of variables can take on with non-zero
probability.
\begin{definition}[Support]
  $\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
  \pmf(x_1 \mapsto v_1, \ldots, x_n \mapsto v_n) > 0 \} $
\end{definition}

>>>>

We define discrete joint probability mass functions (pmfs) in a
standard manner but develop some notations that are useful for our
presentation. Firstly, whereas distributions typically map
realizations of random variables to values in $[0..1]$, for
convenience and following \cite{barthe2019probabilistic} we use
memories to represent realizations, so for example given a pmf $\pmf$
over variables $\{ \sx{x}{1}, \mx{y}{2} \}$ we write $\pmf(\{
\elab{\secret{x}}{1} \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to
denote the (joint) probability that $\elab{\secret{x}}{1} = 0 \wedge
\elab{\mesg{y}}{2} = 1$. Recall from Section
\ref{section-lang-semantics} that $\uplus$ denotes the combination of
memories, so for example $\{ \elab{\secret{x}}{1} \mapsto 0\} \uplus
\{\elab{\mesg{y}}{2} \mapsto 1 \} = \{ \elab{\secret{x}}{1} \mapsto 0,
\elab{\mesg{y}}{2} \mapsto 1 \}$.
\begin{definition}
  A \emph{probability mass function} $\pmf$ is a function
  mapping memories in $\mems(X)$ for given variables $X$ to
  values in $\mathbb{R}$ such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
  %The \emph{support} of a distribution is the set of realizations
  %of random variables with non-zero probability:
  %$\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
  %\pmf(\{ x_1 \mapsto v_1, \ldots, x_n \mapsto v_n\}) > 0 \} $
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ with $\dom(\pmf) = \mems(Y)$, the \emph{marginal distribution}
  of variables $X \subseteq Y$ in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
  $$
  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
  \sum_{\store' \in \mems(Y-X)} \pmf(\store \uplus \store')
  $$
\end{definition}

\begin{definition}
  Given $\pmf$ with $\dom(\pmf) = \mems(Z)$, the \emph{conditional distribution}
  of $X$ given $Y$ where $X \cup Y \subseteq Z$ and $X \cap Y = \varnothing$, denoted $\condd{\pmf}{X}{Y}$,
  is defined as follows:
  $$
  \forall \store \in \mems(X \cup Y)\ .\ 
  \condd{\pmf}{X}{Y}(\store) =
  \begin{cases}
    0 \text{\ if\ } \margd{\pmf}{Y}(\store_Y) = 0\\
    \margd{\pmf}{X \cup Y}(\store) / \margd{\pmf}{Y}(\store_Y)\ \text{\ otherwise}
  \end{cases}
  $$
\end{definition}
We also define some convenient syntactic sugarings. The first will allow us to
compare marginal distributions under different realization conditions
(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
\begin{definition}
  Given $\pmf$, for all $\store_2 \in \mems(Y)$ and $\store_1 \in \mems(X)$ define:
  \begin{enumerate}
  \item $\condd{\pmf}{X}{\store_2}(\store_1) \defeq \condd{\pmf}{X}{Y}(\store_1 \uplus \store_2)$
  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X}(\store_1)$ 
  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X}{Y}(\store_1 \uplus \store_2)$
  \end{enumerate}
\end{definition}
